---
title: "Linear Regression"
author: "Jin Seo Jo"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
```

```{r}
data(stackloss)

fit <- lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc., data = stackloss)

coefficient <- coef(fit) %>% 
  # Access 'enframe' function from the 'tibble' package
  tibble::enframe(name = "term", value = "estimate")

confidence <- confint(fit) %>% 
  as_tibble(rownames = "term")

summary_tibble <- left_join(coefficient, confidence, by = "term")

coefficient
confidence
summary_tibble
```

- `enframe()` converts named atomic vectors or lists to one- or two- column data frames.  
- `as_tibble()` turns an existing object, such as a data frame or matrix, into a so-called tibble, a data frame with class `tbl_df`.  

The script loads the stackloss data, fits a linear regression, extracts the coefficients and confidence intervals and then puts them together in a single object.  

Scripts are excellent for “on the fly” programming, but for repeating task it is useful to wrap this up in a function. So the next task is to make a function called `tidy_lm` with the function signature (aka how it’s called) `tidy_lm(formula, data)`.  

Make a funciton called `tidy_lm` with the function signature `tidy_lm(formula, data)`:
```{r}
tidy_lm <- function(formula, data) {
  
  fit <- lm(formula, data)
  coefficient <- coef(fit) %>% tibble::enframe(name = "term", value = "estimate")
  confidence <- confint(fit) %>% as_tibble(rownames = "term")
  summary_tibble <- left_join(coefficient, confidence, by = "term")
}

data(stackloss)
summary_tibble <- tidy_lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc., stackloss)
```

The function that I have written is possibly not _safe_.  
To see if that's true, restart R (in the Session menu) and try to run that script again. 
Then I get the following error:  
'Error in coef(fit) %>% tibble::enframe(name = "term", value = "estimate") :   
  could not find function "%>%"'  

What went wrong? I didn't import `tidyverse` so the function can't find the pipe.  
There are two options here. The first is to re-write the code so that any packages are explicitly called.  

Re-write the funciton without using a pipe:
```{r}
tidy_lm <- function(formula, data) {
  
  fit <- lm(formula, data)
  coefficient <- tibble::enframe(coef(fit), name = "term", value = "estimate")
  confidence <- as_tibble(confint(fit), rownames = "term")
  summary_tibble <- left_join(coefficient, confidence, by = "term")
}

data(stackloss)
summary_tibble <- tidy_lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc., stackloss)
```

We get the following error:  
'Error in as_tibble(confint(fit), rownames = "term") :  
  could not find function "as_tibble"'   

We need to add some package identifiers (like `tibble::as_tibble`):
```{r}
tidy_lm <- function(formula, data) {
  
  fit <- lm(formula, data)
  coefficient <- tibble::enframe(coef(fit), name = "term", value = "estimate")
  confidence <- tibble::as_tibble(confint(fit), rownames = "term")
  summary_tibble <- dplyr::left_join(coefficient, confidence, by = "term")
}

data(stackloss)
summary_tibble <- tidy_lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc., stackloss)
summary_tibble
```

The second is to check if a required package is loaded and if it is not either load it or print an error message.  

To do this we need `.packages()`. The `.packages()` function returns a vector of strings that name each package that is attached.  
For instance the `stat` package is alwyas attached in an R session (it's a base package with things like `rnorm` in it).
```{r}
"stats" %in% .packages()
```

What if a package isn't installed?  
We can use `require(dplyr)` instead of `library(dplyr)`. The difference is that `require` returns a logical values (TRUE/FALSE) depending on if the package is available, whereas `library` will just throw an error. 

Check if `dplyr` and `tibble` are attached and, if they are not, use `stop()` to send a useful error message. 
```{r}
if(!all(c("dplyr", "tibble") %in% .packages())) {
    stop("You must have the dplyr and tibble packages attached!")
}
```

Check if `dplyr` and `tibble` are attached and, if they are not attach them and add it to the top of the function. 
```{r}
if(!all(c("dplyr", "tibble") %in% .packages())) {
  library(dplyr)
  library(tibble)
  # install.packages("dplyr")
}
```

Use `require` to attach a package if it is installed and throw a useful error message if it is not.  
```{r}
if(!require(dplyr)) {
  stop("The dplyr packages must be installed. Run install.packages(\"dplyr\") and then try again.")
}
if(!require(tibble)) {
  stop("The tibble packages must be installed. Run install.packages(\"dplyr\") and then try again.")
}
```

Finally, no funciton is complete without some documentation! Good function documentation should  
- Describe what it does
- Describe what goes in 
- Describe what comes out
- Give a quick example of how it works

A skeleton is here:
```{r}
my_function <- function(a, b = "2009"){
## my_function computes something
## Example my_function computes the death and birth rates in Canadian provinces ## in a given year
##
## Input:
## - a: A [type] that [what should it mean]. [If there is something that needs
  ##      to be true, say it here]. Example: A character vector of two-letter
  ##      Province abbreviations.
  ## - b: (Optional) A [type] that [what should it mean]. Default = "2009".
  ##   Example: The year as a string. Any year between 2000 and 2015
  ##
  ## Output:
  ## - Returns a [type] that [describe how to interpret the return]. Example:
  ##   Returns a list of birth and death rates for the provinces in a in year b.
  ##   The first element is the birth rate in year b, the second element is the
  ##   death rate in 2009.
  ##
  ## - Example:
  ## rates <- my_function(c("ON", "NB"), "2010")
  # Function code goes here
}
```

If we are building an R package instead of just documenting a loose funciton, we should use `Roxygen2`. The major difference is the specific formatting, but it lets us automatically generate package documentaion!

Now we write documentation for the `tidy_lm` function. 
```{r}
tidy_lm <- function(formula, data) {
  ## tidy_lm performs the linear regression lm(formula, data) and then
  ## collects the estimates and the confidence interval in a single tibble.
  ##
  ## Input:
  ## - formula: A formula object for the linear regression
  ## - data: Data for the linear regression
  ##
  ## Output:
  ## - A tibble that has columns for the estimate. The 2.5% confidence boundary
  ## 97.5% confidence boundary. Each row is one term in the formula.
  ##
  ## Example:
  ## data(stackloss)
  ## summary <- tidy_lm(stack.loss ~ ., stackloss)
  
  if(!require(dplyr)) {
    stop("The dplyr packages must be installed. Run install.packages(\"dplyr\") and then try again.")
  }
  if(!require(tibble)) {
    stop("The tibble packages must be installed. Run install.packages(\"dplyr\") and then try again.")
  }
  
  fit <- lm(formula, data)
  coefficient <- tibble::enframe(coef(fit), name = "term", value = "estimate")
  confidence <- tibble::as_tibble(confint(fit), rownames = "term")
  summary_tibble <- dplyr::left_join(coefficient, confidence, by = "term")
}
```

Finally, this exercise was a partial reimplementation of the funtion `broom:::tidy.lm`. We should look at the code for `broom:::tidy.lm` and the output of `tidy(fit, conf.int = TRUE)` to compare out code with some professional R code.  

Use the data set `mtcars` to compare the broom implementation and our implementation.
```{r}
data(mtcars)
my_tidy <- tidy_lm(mpg ~ ., mtcars)
my_tidy
```

```{r}
fit <- lm(mpg ~ ., mtcars)
their_tidy <- broom::tidy(fit, conf.int = TRUE)
their_tidy
```

_A notte on naming conventions in R_: Why is that function called `broom:::tidy.lm`? Firstly, the three colons says that this function isn't directly exported for uses use. The broom package exports the function `tidy` instead. It has very simple code
```{r}
broom::tidy
```

The function `UseMethod("tidy")` basically tells R to do the following two steps:
1. Work out what class `x` by calling `class(x)`. (`fit` has class `lm`)
2. Find a fucntion called `tidy.[class(x)]` and execute that. In this case it finds `tidy.lm()` deep inside the `broom` package and calls that.  

This is a trick that is used frequently in R programming to make sure that generic functions (like `tidy` or `summary`) can work across a bunch of different types of inputs. 
